### 2014-10-25 motivation.tex
Edits to motivation

\name automates this error-prone task by letting the programmer
declare application-level consistency constraints as \emph{contracts}.
This obviates the need to understand store-level consistency guarantees,
and relieves the programmer from having to match application
requirements with appropriate store-level alternatives. The
cornerstone of declarative reasoning in \name is the programming model
that draws largely on the concept of operation-based
CRDTs~\cite{shapiroCRDT}. The state of a replicated data object in
\name is nothing but the set of effects applied on the object. For
example, to represent a \cf{BankAccount} object in \name, we first
define an sum type of all possible effects on a \cf{BankAccount}
object:
%types in Quelea are defined as reductions over the set of effects.
%The following data type definitions capture the operations and effects
%on the bank account and user index objects.
\begin{codehaskell}
type Amount = Float
data BankAccountEff = Deposit Amount
                    | Withdraw Amount 
                    | GetBalance
\end{codehaskell}
Subsequently, any \cf{BankAccount} object can be simply represented as a set of
effects of type \cf{BankAccountEff} that took effect on the object.
For example, a \cf{BankAccount} object obtained by performing a
deposit of 100 and a withdrawl of 50 on a fresh bank account is
represented as following:
\begin{center}
  \{\cf{Deposit(100), Withdraw(50)}\}
\end{center}
An operation on a replicated data object in \name is simply defined as
a reduction over the set of effects representing the object, and is
expected to produce a new effect along with a return value. For
instance, following are the definitions of operations on
\cf{BankAccount} type in \name (we represent sets as lists for
clarity):
\GK{Better define getBalance in terms of fold. It goes well with our
narrative.}
\begin{codehaskell}
type BankAccount = [BankAccountEff]
getBalance ::  BankAccount {- object -}
  -> () {- args -}
  -> (Amount {- ret val -}, 
      Maybe BankAccountEff {- effect -})
getBalance ctxt _ = (sum [v | Deposit v (- ctxt]
        - sum [v | Withdraw v (- ctxt], Nothing)

@@

\GK{It is better to split this contract into two parts.}
If a withdraw $b$ is visible to the get balance operation, then all
deposit operations $a$ visible to withdraw should be visible to the
get balance operation. This prevents negative balance anomalies. The
rest of the contract says that a get balance operation must witness
previous deposit and withdraw operations on the same object in the
same session. This prevents missed update anomalies.

>>>>

\name automates this error-prone task by letting the programmer
declare application-level consistency constraints as \emph{contracts}.
% This obviates the need to understand store-level consistency guarantees,
% and relieves the programmer from having to match application
% requirements with appropriate store-level alternatives. 
The cornerstone of declarative reasoning in \name is the programming
model that draws largely on the concept of operation-based
CRDTs~\cite{shapiroCRDT}. The state of a replicated data object in
\name is nothing but the set of effects applied on the object. For
example, to represent a \cf{BankAccount} object in \name, we first
define an sum type of all possible effects on a \cf{BankAccount}
object:
%types in Quelea are defined as reductions over the set of effects.
%The following data type definitions capture the operations and effects
%on the bank account and user index objects.
\begin{codehaskell}
type Amount = Float
data BankAccountEff = Deposit Amount
                    | Withdraw Amount 
                    | GetBalance
\end{codehaskell}
Subsequently, any \cf{BankAccount} object can be simply represented as a set of
effects of type \cf{BankAccountEff} that took effect on the object.
For example, a \cf{BankAccount} object obtained by performing a
deposit of 100 and a withdrawl of 50 on a fresh bank account is
represented as following:
\begin{center}
  \{\cf{Deposit(100), Withdraw(50)}\}
\end{center}
An operation on a replicated data object in \name is simply defined as
a reduction over the set of effects representing the object, and is
expected to produce a new effect along with a return value. For
instance, following are the definitions of operations on
\cf{BankAccount} type in \name (we represent sets as lists for
clarity):
\GK{I think it is better define getBalance in terms of fold. It goes
well with our narrative.}
\begin{codehaskell}
type BankAccount = [BankAccountEff]
getBalance ::  BankAccount {- object -}
  -> () {- args -}
  -> (Amount {- ret val -}, 
      Maybe BankAccountEff {- effect -})
getBalance ctxt _ = (sum [v | Deposit v (- ctxt]
        - sum [v | Withdraw v (- ctxt], Nothing)

@@

\GK{It may be better to split this contract into two parts.}
If a withdraw $b$ is visible to the get balance operation, then all
deposit operations $a$ visible to withdraw should be visible to the
get balance operation. This prevents negative balance anomalies. The
rest of the contract says that a get balance operation must witness
previous deposit and withdraw operations on the same object in the
same session. This prevents missed update anomalies.
